Hydraulically-vulnerable trees survive on deep-water access in a tropical forest community

### 2020-10-23 manuscript.rmd
lifespan replaced with lifetime

#### Leaf Area Index (LAI)
\
In the exploratory model for effective rooting depth, we explored the effect of seasonality in leaf area index (LAI, unitless) on growth. We assumed species-specific mean seasonal curves for LAI normalised between 0 and 1 (LAI, unitless), informed by a combination of long-term records for leaf-fall [@Wright.1990aeo] and leaf-lifespan [@Osnas.2018] (see SI).

@@

We generated an estimate of the mean seasonality curve for normalised leaf area index (LAI$^*$, range 0-1, unitless) for individual species as follows. We interpolated species-specific weekly leaf-fall data collected at two similar sites in BCI (1985-2020) to a daily estimate [@Wright.1990aeo]. We lead this daily leaf-fall time series by the species-specific average leaf-lifespan (in days) and used it as an estimate of the daily leaf-gain time-series. The difference between leaf-fall and leaf-gain gave an estimate of daily leaf-change time series, which was then scaled to express as a fraction of the total leaf fall in a given year. Beginning with the LAI$^*$ of 1 at the end of the wet season in the first year of the record, we acccumulated leaf-change over it to obtain a running estimate of the leaf-cover fraction, avaraged it by DOY across years and normalised it to obtain the species-specific seasonality curve in LAI$^*$ (range 0-1, unitless). (Fig. S\@ref(fig:leaf-cover)).

For five species, an estimate of leaf lifespan was available from direct observations [@Osnas.2018]. For the rest erd.sp.wo.ll species, leaf lifespan was predicted from LMA ($\overline{lifespan} = `r ll.lma.cf[1]` + `r ll.lma.cf[2]`LMA$; *R*^2^ = `r ll.lma.r2`; *p* `r ll.lma.p`, *n* = `r ll.lma.n`).

>>>>

#### Leaf Area Index (LAI)
\
In the exploratory model for effective rooting depth, we explored the effect of seasonality in leaf area index (LAI, unitless) on growth. We assumed species-specific mean seasonal curves for LAI normalised between 0 and 1 (LAI, unitless), informed by a combination of long-term records for leaf-fall [@Wright.1990aeo] and leaf-lifetime [@Osnas.2018] (see SI).

@@

We generated an estimate of the mean seasonality curve for normalised leaf area index (LAI$^*$, range 0-1, unitless) for individual species as follows. We interpolated species-specific weekly leaf-fall data collected at two similar sites in BCI (1985-2020) to a daily estimate [@Wright.1990aeo]. We lead this daily leaf-fall time series by the species-specific average leaf-lifetime (in days) and used it as an estimate of the daily leaf-gain time-series. The difference between leaf-fall and leaf-gain gave an estimate of daily leaf-change time series, which was then scaled to express as a fraction of the total leaf fall in a given year. Beginning with the LAI$^*$ of 1 at the end of the wet season in the first year of the record, we acccumulated leaf-change over it to obtain a running estimate of the leaf-cover fraction, avaraged it by DOY across years and normalised it to obtain the species-specific seasonality curve in LAI$^*$ (range 0-1, unitless). (Fig. S\@ref(fig:leaf-cover)).

For five species, an estimate of leaf lifetime was available from direct observations [@Osnas.2018]. For the rest erd.sp.wo.ll species, leaf lifetime was predicted from LMA ($\overline{lifetime} = `r ll.lma.cf[1]` + `r ll.lma.cf[2]`LMA$; *R*^2^ = `r ll.lma.r2`; *p* `r ll.lma.p`, *n* = `r ll.lma.n`).
